---
title: "Download Savant Function"
author: "Liam Jennings"
date: "2025-03-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PURPOSE: Initialize utility function from Dr. Scott Powers for gathering pitch level data

Download Statcast data from Baseball Savant
 
Loop over five days at a time to download data from the Statcast search API at Baseball Savant.

This includes swing tracking data not available through the MLB statsapi.

<br>

**@param start_date first date included in the download**
**@param end_date last date included in the download**
**@param verbose logical, should progress be printed?, defaults to FALSE**

<br>

**@return a dataframe with 113 columns (as of now) and one row per pitch, with all available data**


### Function
```{r}
download_savant_data <- function(start_date, end_date, verbose = FALSE) {
  
  base_url <- "https://baseballsavant.mlb.com/statcast_search/csv?all=true&type=details"
  
  # Split the dates into 5-day chunks. The Savant API will return at most 25,000 rows. Assuming 300
  # pitches per game, a day with 15 games will have 4,500 pitches. We can safely download 5 days of
  # data, but more days would risk hitting the 25,000-row limit.
  days <- as.numeric(as.Date(end_date) - as.Date(start_date))
  start_date_seq <- as.Date(start_date) + seq(from = 0, by = 5, to = days)
  end_date_seq <- start_date_seq + 4
  end_date_seq[length(end_date_seq)] <- end_date
  
  data <- NULL
  for (i in 1:length(start_date_seq)) {
    
    if (verbose) {
      message(glue::glue("Downloading Savant data from {start_date_seq[i]} to {end_date_seq[i]}"))
    }
    
    url <- glue::glue("{base_url}&game_date_gt={start_date_seq[i]}&game_date_lt={end_date_seq[i]}")
    data_i <- read.csv(url(url))
    
    if (nrow(data_i) == 25000) {
      warning(
        glue::glue("Exactly 25,000 rows returned for {start_date_seq[i]} to {end_date_seq[i]}")
      )
    }
    
    data <- dplyr::bind_rows(data, data_i)
  }
  
  return(data)
}
```

### Load Statcast Data
```{r}
## 2021
statcast_2021 <- download_savant_data("2021-04-01", "2021-10-03")

## 2022
statcast_2022 <- download_savant_data("2022-04-07", "2022-10-05")

## 2023
statcast_2023 <- download_savant_data("2023-03-30", "2023-10-01")

## 2024
statcast_2024 <- download_savant_data("2024-03-20", "2024-09-30")

## rowbind
statcast <- rbind(statcast_2021, statcast_2022, statcast_2023, statcast_2024)
```


### Save the data
```{r}
write.csv(statcast, "statcast.csv")
```

